\bf Data Assimilation for Modeling and Calibrating of Disease-altered Potassium Channel Kinetics in Mouse Cardiomyocytes

### 2022-08-24 writing/draft.tex
Draft revised

\begin{abstract}
Activities of cardiac voltage-gated potassium channels (K\textsubscript{v}) are essential in the electrical conduction system of the heart, and even slight changes in function can cause fatal heart disease such as long QT syndrome. Our recent studies found that altered glycosylation causes aberrant electrical signaling in mouse cardiomyocytes and leads to heart failure and early death. Activities of K\textsubscript{v} channels were also affected, and K\textsuperscript{+} currents were significantly reduced. Among major cardiac voltage-gated ion channels (VGIC), K\textsubscript{v} channels have a distinctive feature: various isoforms generate their own unique currents contributing to different phases of repolarization of cardiac action potential (AP). On the other hand, the dominant isoform is responsible for Na\textsuperscript{+} and Ca\textsuperscript{2+} currents. This characteristic raises challenges for studying K\textsubscript{v} channels. Although it is highly desirable to observe the activities of each K\textsubscript{v} isoform separately, only collective outputs can be measured because of their slightly overlapping voltage-dependence of gating. In whole-cell recordings, the sum of K\textsuperscript{+} currents (I\textsubscript{Ksum}) can be measured. Traditional methods apply curve fitting to estimate components of K\textsubscript{v} isoforms from I\textsubscript{Ksum}. Additional curve fitting is used to find kinetics information for the individual K\textsuperscript{+} currents. However, this traditional approach is limited in its ability to determine gating kinetics of K\textsubscript{v} isoforms rigorously. Curve fittings do not account for interactions of multiple gating processes of various isoforms and dynamics according to changes in protocols. Therefore, we propose a novel data assimilation framework for investigating whole-cell K\textsubscript{v} current traces. First, computer models of K\textsuperscript{+} currents are designed with parameters that control the kinetic rates and, in turn, determine the currents. Second, we develop fractional factorial designs to identify the parameters that have significant impacts on the current shapes. Third, we develop model calibration as nonlinear optimization that couple the \textit{in-silico} K\textsubscript{v} models and \textit{in-vitro} I\textsubscript{K} recordings. We apply the proposed framework to our  glycosylation data. Experimental results show that the data assimilation method can estimate differences in kinetic rate between the control and experiment groups. This method has strong potential to pave a new way to analyze K\textsubscript{v} and heart diseases.
\end{abstract}

\noindent%
{\it Keywords:} Data assimilation, potassium channels, kinetics modeling, model calibration, glycosylation

@@

The main advantage of this approach is that it analyzes I\textsubscript{K} recordings based on the biophysical models with considering cellular level dynamics according to changes in protocols for studying heart diseases. Our contributions are summarized as follows:
\begin{itemize}
    \item We modeled the underlying kinetics of K\textsubscript{v} isoforms from whole-cell I\textsubscript{K} recordings based on biophysical principles rather than curve-fitting methods.
    \item We developed a model calibration routine that harnesses multiple protocols simultaneously rather than a single protocol independently, resulting in comprehensive cellular-level modeling.
    \item cell variability.
    \item Our method captures differences in channel kinetics given the disease-related perturbation, such as reduced glycosylation.
\end{itemize}
The proposed data assimilation framework has strong potential to serve as a novel complementary data analytics tool to in-vitro experiments.

\section{Research Background}
\subsection{Computational Modeling of Cardiomyocytes}
computational electrophysiology (pain points in analyziyng or modeling the data with focus on Kv). add citation \citep{chen2019heterogeneous}

Computational approaches for studying heart (cite papers of Bing or Rui). Narrow down the topics into molecular-level studies: action potential and iom channels. Computational electrophyisiology has been a successful story of system biology. It has provided new insights how excitable cells work (neurons, cardiomyocytes, etc.). It can be also used to test new drugs. Other many application are also available. It is one of the vital part of disease-related investigation.

Example: studies of the effects of altered glycosylation in the heart and pain points (elaborate the pain points mentioned in the earlier paragraph)

calibration is necessary to apply this computational approaches (or models). How the calibration has been done in the literature.

Point out problems of curve fitting

\subsection{Altered Glycosylation and Heart Diseases}
DCM and altered glycosylation

@@

We developed the box-constrained nonlinear optimization routine using nonlinear optimization algorithms with multi-random initial points to minimize $\delta$. Box constraints mean that $\theta$ has a lower and upper bound for each dimension, so solution space is constrained in a hypercube. In this way, the optimization loop can be controlled by users, allowing them to blend their domain knowledge into the modeling. Note that it is possible because the suggested method is based on biophysical models that provide interpretability of $\theta$ in channel kinetics rather than the curve fitting. Global optimum is not guaranteed due to the complexity of our models. We developed the multi-random starting scheme to avoid local optima and find the solution as close to the optimum as possible. Latin hypercube designs and parallel computations are used to sample initial points and run them on multicores to compensate for the increased computational burden. Different nonlinear optimization algorithms are tested and compared. This work is implemented in MATLAB R2022a, and the code is available at xyz.

@@

\section{Experimental Design}
We applied the proposed framework to our investigations in the pathophysiology of electrical signaling by altered glycosylation. Protein glycosylation is one of the most abundant and diverse forms of co/posttranslational modifications that impact essential protein functions, such as modulation of receptor or ion channel activities \citep{ohtsubo2006glycosylation,ednie2012modulation}. A growing number of studies has shown the association between altered glycosylation and heart diseases, such as dilated cardiomyopathy (DCM) and hypertrophic cardiomyopathy \citep{ohtsubo2006glycosylation,ednie2019reduced2}. It is reported that up to 20\% of patients with congenital disorders of glycosylation (CDG), who commonly have modest reductions in protein glycosylation, present with cardiac deficits and idiopathic DCM \citep{marques2017cardiac}. We have investigated how regulated glycosylation contributes to heart failure in the context of electrophysiology. Electrical signaling is orchestrated activities of a variety of ion channels and transporters. Voltage-gated ion channels (VGICs) are heavily glycosylated, with ~30\% of the channel mass consisting of N-/O-linked glycans \citep{ednie2012modulation}.

Glycosylation is a multi-step process and usually ends with sialic acid added. We reported that a saturating, electrostatic effect of negatively charged sialic attached to the terminal of N-/O-glycan branches significantly altered electrical signaling in Na\textsubscript{v} \citep{ednie2013sialicNav1,ednie2015sialicNav2} as well as K\textsubscript{v} \citep{ednie2015sialicKv}. More recently, we showed that preventing hybrid/complex N-glycosylation in mouse cardiomyocytes was sufficient to cause DCM using genetic ablation of the \textit{MGAT1} gene (MGAT1KO model), which encodes a critical glycosyltransferase, GlcNAcT1 \citep{ednie2019reduced2}. MGAT1KO mice developed DCM, deteriorated into heart failure, and 100\% died early, presumably from ventricular arrhythmias leading to sudden cardiac death. To further investigate the role of altered glycosylation and pathogenesis and disease progression of the heart, we conducted whole-cell patch-clamp experiments that showed reductions in N-glycosylation significantly impact electrical signaling in mouse cardiomyocytes \citep{ednie2019reduced}. To be specific, whole-cell I\textsubscript{K} traces were measured in left ventricular apex cardiomyocytes of ~14 week-old control (wild type; WT) and MGAT1KO mice, elicited by 4.5 s 10mV voltage steps (-30 - 50 mV) from holding potential of -70 mV. There were 31 and 30 observations for each WT and MGAT1KO.

It was observed in the experiment that K\textsuperscript{+} currents were reduced in MGAT1 ventricular myocytes. A curve-fitting method of bi-exponential, combining I\textsubscript{Kslow1} and I\textsubscript{Kslow2} of tri-exponential in Equation~\ref{eq:expl_fitting}, was applied to decompose I\textsubscript{Ksum} traces into component currents. All three component currents were reduced, but the rectifier current was significantly reduced and slowed notably. Although the \textit{in-vitro} investigation discovered aberrant reductions in K\textsuperscript{+} currents with altered glycosylation, it was difficult to determine channel kinetics rigorously. For example, extra curve fitting was not available due to small current magnitude of the MGAT1KO model. Therefore, we leverage the suggested framework for modeling K\textsubscript{v} isoforms kinetics from our experimental data of I\textsubscript{K} recordings. This new approach allows dissecting whole-cell K\textsuperscript{+} current traces into isoform components and modeling their underlying kinetics with considering the cellular-level dynamics of multiple protocols.

\section{Results}
\subsection{Parameter Screening}
\textbf{Figure~\ref{fig:half-norm}} shows the half-normal plots of factorial effects of parameters on the six marker points. The red straight lines on the plots serve as a criterion to identify the parameters that are sensitive to the marker points. The farther parameters fall above the straight lines, the more significant impacts they have on the marker. We picked the parameters to be calibrated as a union of parameter sets above the straight lines on the six half-normal plots. For I\textsubscript{Kto} $\{p_{1},\ p_{2},\ p_{3},\ p_{4},\ p_{5},\ p_{7},\ p_{11}\}$ are selected, for I\textsubscript{Kslow1} $\{p_{1},\ p_{2},\ p_{4},\ p_{5},\ p_{9},\ p_{10}\}$, for I\textsubscript{Kslow2} $\{p_{1},\ p_{2}\}$, and for I\textsubscript{Kss} $\{p_{1},\ p_{2},\ p_{3}\}$. 
\begin{figure}
    \centering
    \includegraphics{figs/half-norm.pdf}
    \caption{Half-normal plots of factorial effects of parameters on the six maker points.}
    \label{fig:half-norm}
\end{figure}

\textbf{Figure~\ref{fig:calib_param}} shows the selected parameters, highlighted in different colors according to their functional roles in channel kinetics. We categorized the calibration parameters into four classes: The red represents the voltage-threshold parameters and the green voltage slopes, controlling the voltage dependence, the blue scale factors of kinetic functions, and purple time-constant shifters. Note that the voltage-dependence parameters in red and green appear multiple times across different equations. This parsimonious model design is intended to maximize the structural regularization to minimize overfitting.
\begin{figure}[!ht]
    \centering
    \includegraphics{figs/calib_param.pdf}
    \caption{Selected parameters from the sensitivity analysis, highlighted in different colors according to their functional roles in channel kinetics.}
    \label{fig:calib_param}
\end{figure}

\subsection{Model Fitness}
We compared three nonlinear optimization algorithms satisfying box constraints: BFGS, sequential quadratic programming (SQP), and active set methods. \textbf{Fig~\ref{fig:model_fitness}A} and \textbf{C} shows values of the objective function (sum of RMSEs) of the three optimization algorithms for WT and MGAT1KO, respectively. The three algorithms barely showed differences in the objective values. We compared the actual model predictions and experimental potassium I\textsubscript{K} traces in \textbf{Figure~\ref{fig:model_fitness}B} and \textbf{D} to validate the model calibration results further. Representative data were selected randomly for each group and plotted at -30 mV and 50 mV voltage steps to illustrate discrepancies at low and high voltages. Sum of RMSEs values are lower in MGAT1KO than WT in general, because magnitudes are smaller for the disease group. Calibration results of the BFGS algorithm will be used in analysis of kinetics modeling in the rest the sections. 
\begin{figure}[!ht]
    \centering
    \includegraphics{figs/model_fitness_4half4.pdf}
    \caption{Bar graphs of the sum of RMSEs of (A) WT and (C) MGAT1KO, and representative actual fitness between model outputs and experimental I\textsubscript{K} recordings (in red) at -30 mV and 50 mV voltage steps of (B) observation 26 in WT and (D) 4 in MGAT1KO.}
    \label{fig:model_fitness}
\end{figure}

>>>>

\begin{abstract}
Activities of cardiac voltage-gated potassium channels (K\textsubscript{v}) are essential in the electrical conduction system of the heart. Even slight changes in K\textsubscript{v} functions can cause fatal heart disease, such as long QT syndrome. Our recent studies found that altered glycosylation causes aberrant electrical signaling in mouse cardiomyocytes and leads to heart failure and early death. Activities of K\textsubscript{v} channels were also affected by altered glycosylation, and K\textsuperscript{+} currents were significantly reduced. Among major cardiac voltage-gated ion channels (VGIC), K\textsubscript{v} channels have a distinctive feature: various isoforms generate their own unique currents contributing to different phases of repolarization of the cardiac action potential (AP). On the other hand, the dominant isoform is responsible for Na\textsuperscript{+} and Ca\textsuperscript{2+} currents. This characteristic raises challenges for studying K\textsubscript{v} channels. Although it is highly desirable to observe the activities of each K\textsubscript{v} isoform separately, only collective outputs can be measured because of their slightly overlapping voltage-dependence of gating. In whole-cell recordings, the sum of K\textsuperscript{+} currents (I\textsubscript{Ksum}) can be measured. Traditional methods apply curve fitting to estimate components of K\textsubscript{v} isoforms from I\textsubscript{Ksum}. Additional curve fitting is used to find kinetics information for the individual K\textsuperscript{+} currents. However, this traditional approach is limited in its ability to determine gating kinetics of K\textsubscript{v} isoforms rigorously. Curve fittings do not account for interactions of multiple gating processes of various isoforms and dynamics according to changes in protocols. Therefore, we propose a novel data assimilation framework for investigating whole-cell K\textsubscript{v} current traces and apply the proposed method to our glycosylation data. First, computer models of K\textsuperscript{+} currents are designed with parameters that control the kinetic rates and, in turn, determine the currents. Second, we develop fractional factorial designs to identify the parameters that have significant impacts on the current shapes. Third, we develop model calibration as nonlinear optimization that couple the \textit{in-silico} K\textsubscript{v} models and \textit{in-vitro} I\textsubscript{K} recordings. Last, low-dimensional embedding is used to visualize distributional difference of kinetic parameters of cells, showing the cellular-level variability. Experimental results show that the data assimilation method can estimate differences in kinetic rate between the control and experiment groups. This method has strong potential to pave a new way to analyze K\textsubscript{v} and heart diseases.
\end{abstract}

\noindent%
{\it Keywords:} Data assimilation, model calibration, potassium channels, kinetics modeling, glycosylation

@@

The main advantage of this approach is that it analyzes I\textsubscript{K} recordings based on the biophysical models with considering cellular level dynamics according to changes in protocols for studying heart diseases. Our contributions are summarized as follows:
\begin{itemize}
    \item We modeled the underlying kinetics of K\textsubscript{v} isoforms from whole-cell I\textsubscript{K} recordings based on biophysical principles rather than curve-fitting methods.
    \item We developed a model calibration routine that harnesses multiple protocols simultaneously rather than a single protocol independently, resulting in comprehensive cellular-level modeling.
    \item The proposed approach builds separate K\textsubscript{v} models and calibrates them for each cell, which allows analyzing the cellular variability. 
    \item Our method captures differences in channel kinetics given the disease-related perturbation, such as reduced glycosylation.
\end{itemize}
The proposed data assimilation framework has strong potential to serve as a novel complementary data analytics tool to in-vitro experiments.

\section{Research Background}
\subsection{Computational Modeling of the Heart}
Computer models of the heart have been extensively studied and used in cardiology to understand the mechanism of the organ, investigate fatal diseases, and develop drugs. Models consist of mathematical equations describing biophysical properties and physiological functions compatible with experimental observations. Mathematical and computational implementations of the heart allow studying the heart in a quantitative and predictive way \citep{whittaker2020calibration}. For example, models can be used to interpolate processes not directly observed in experiments and extrapolate to novel conditions such as disease-related perturbations \citep{rodriguez2010systems}. With the advancement of sensing technology and data analysis, computational cardiology is expanding its field. Personalized heart models pave a new way for monitoring and managing the cardiac health of individuals by characterizing individual heart signals from the large-scale IoT network and calculating dissimilarities to detect potential cardiac risk \citep{yang2020network}. Hence, cardiac computer models provide a great opportunity to mitigate heart disease, the number one cause of death in the U.S. \citep{murphy2021mortality}.

The functionality of the heart to pump blood through the body is controlled by electrical signals generated in the heart itself. Hence, the electrical activities and waves spreading through the whole heart have been studied considerably. For examle, \citet{yang2016whole} a new reaction-diffusion and pattern recognition method to model spatiotemporal dynamics of electrical waves propagating throughout the heart surface. Sensing methods, such as vectorcardiogram (VCG) and electrocardiogram (ECG), measure the electrical activities of the heart. Various analytics methods have been proposed to cope with their complex nonlinear dynamics in different heart regions over time. \citet{chen2019heterogeneous} developed a heterogeneous recurrence method that characterizes spatiotemporal patterns in disease-altered multi-channel cardiac signals. Nonlinear recurrence analysis captures and quantifies topological relationships of recurring patterns in signals, which has been proven to be an effective tool for cardiac signals that are naturally recurring over time \citep{yang2010multiscale,meyers2020cross}.

However, to perform a fine-grained analysis, it is necessary to study the cellular-level activities of the heart as opposed to the organ-level signals. Voltage-gated ion channels (VGICs) are membrane proteins in cardiomyocytes that are the functional unit of the electrical activities of cardiac muscle cells. The whole-cell patch clamp recording is a powerful experimental technique for investigating the electrical properties of excitable cells such as cardiomyocytes and neurons. It allows measuring ionic currents conducted through designated VGICs and the action potential (AP). Although the laboratory technique reveals essential properties of VGICs and cardiomyocytes, the detailed gating mechanism remains ambiguous. For example, it is hard to determine rigorously how different ion channels collectively lead to changes in the AP. Computer models, coupled with experimental observations, provide integrative insights into the data, and they have become an essential part of studying the heart, as illustrated in Figure~\ref{fig:framework_diagram} \citep{winslow2011integrative}.

Hence, modeling VGICs involves a data assimilation procedure that combines experimental data with computer models. Traditionally, computer models are calibrated to characteristic curves estimated from ionic-current recordings \citep{colman2018description}. These estimation processes include curve fitting, which assumes a functional form of the characteristic curve and finds shape parameters best fit data. However, curve fitting only considers a single current recording at each time independently, so it cannot capture cellular-level dynamics from the current recordings by changes in input stimuli, i.e., voltage-clamp protocols. It raises an additional challenge for analyzing I\textsubscript{K} recordings, because one-more curve fitting is used to separate individual current traces generated by various K\textsubscript{v} isoforms. I\textsubscript{Ksum} traces are decomposed only based on the assumption of the shapes of the component currents but not the underlying biophysical gating kinetics \citep{kim2022simulation}. Therefore, there is an urgent need to develop a novel data assimilation method for modeling and calibrating K\textsubscript{v} kinetics with considering cellular dynamics.

\subsection{Altered Glycosylation and Heart Diseases}
Protein glycosylation is one of the most abundant and diverse forms of co/posttranslational modifications that impact essential protein functions, such as modulation of receptor or ion channel activities \citep{ohtsubo2006glycosylation,ednie2012modulation}. A growing number of studies has shown the association between altered glycosylation and heart diseases, such as dilated cardiomyopathy (DCM) and hypertrophic cardiomyopathy \citep{ohtsubo2006glycosylation,ednie2019reduced2}. It is reported that up to 20\% of patients with congenital disorders of glycosylation (CDG), who commonly have modest reductions in protein glycosylation, present with cardiac deficits and idiopathic DCM \citep{marques2017cardiac}. However, the underlying pathological mechanism still remains elusive.

We have investigated how regulated glycosylation contributes to heart failure in the context of electrophysiology. Electrical signaling is orchestrated activities of a variety of ion channels and transporters. Voltage-gated ion channels (VGICs) are heavily glycosylated, with ~30\% of the channel mass consisting of N-/O-linked glycans \citep{ednie2012modulation}. Glycosylation is a multi-step process and usually ends with sialic acid added. We also reported that a saturating, electrostatic effect of negatively charged sialic attached to the terminal of N-/O-glycan branches significantly altered electrical signaling in Na\textsubscript{v} \citep{ednie2013sialicNav1,ednie2015sialicNav2} as well as K\textsubscript{v} \citep{ednie2015sialicKv}. Computational modeling has been used to further investigate the functional role of reduced sialylation in Na\textsubscript{v} and K\textsubscript{v} \citep{du2015statistical, du2017silico}. However, these studies also rely on characteristic curves and statistics estimated from curve fitting.

@@

We developed the box-constrained nonlinear optimization routine using nonlinear optimization algorithms with multi-random initial points to minimize $\delta$. Box constraints mean that $\theta$ has a lower and upper bound for each dimension, so solution space is constrained in a hypercube. In this way, the optimization loop can be controlled by users, allowing them to blend their domain knowledge into the modeling. Note that it is possible because the suggested method is based on biophysical models that provide interpretability of $\theta$ in channel kinetics rather than the curve fitting. Global optimum is not guaranteed due to the complexity of our models. We developed the multi-random-start scheme to avoid local optima and find the solution as close to the optimum as possible. Latin hypercube designs and parallel computations are used to sample initial points and run them on multicores to compensate for the increased computational burden. This work is implemented in MATLAB R2022a, and the code is available at xyz.

@@

\section{Experimental Design}
More recently, we showed that preventing hybrid/complex N-glycosylation in mouse cardiomyocytes was sufficient to cause DCM using genetic ablation of the \textit{MGAT1} gene (MGAT1KO model), which encodes a critical glycosyltransferase, GlcNAcT1 \citep{ednie2019reduced2}. MGAT1KO mice developed DCM, deteriorated into heart failure, and 100\% died early, presumably from ventricular arrhythmias leading to sudden cardiac death. To further investigate the role of altered glycosylation in pathogenesis and disease progression of the heart, we conducted whole-cell patch-clamp experiments that showed reductions in N-glycosylation significantly impact electrical signaling in mouse cardiomyocytes \citep{ednie2019reduced}. To be specific, whole-cell I\textsubscript{K} traces were measured in left ventricular apex cardiomyocytes of ~14 week-old control (wild type; WT) and MGAT1KO mice, elicited by 4.5 s 10mV voltage steps (-30 - 50 mV) from holding potential of -70 mV. There were 31 and 30 observations for each WT and MGAT1KO.

We applied the proposed framework to our investigations in the pathophysiology of electrical signaling by altered glycosylation. It was observed in the experiment that K\textsuperscript{+} currents were reduced in MGAT1 ventricular myocytes. A curve-fitting method of bi-exponential, combining I\textsubscript{Kslow1} and I\textsubscript{Kslow2} of tri-exponential in Equation~\ref{eq:expl_fitting}, was applied to decompose I\textsubscript{Ksum} traces into component currents. All three component currents were reduced, but the rectifier current was significantly reduced and slowed notably. Although the \textit{in-vitro} investigation discovered aberrant reductions in K\textsuperscript{+} currents with altered glycosylation, it was difficult to determine channel kinetics rigorously. For example, extra curve fitting was not available due to small current magnitude of the MGAT1KO model. Therefore, we leverage the suggested framework for modeling K\textsubscript{v} isoforms kinetics from our experimental data of I\textsubscript{K} recordings. This new approach allows dissecting whole-cell K\textsuperscript{+} current traces into isoform components and modeling their underlying kinetics with considering the cellular-level dynamics of multiple protocols.

\section{Results}
\subsection{Parameter Screening}
\textbf{Figure~\ref{fig:half-norm}} shows the half-normal plots of factorial effects of parameters on the six markers. The red straight lines on the plots serve as a criterion to identify the parameters that are sensitive to the marker points. The farther parameters fall above the straight lines, the more significant impacts they have on the marker. We picked the parameters to be calibrated as a union of the sets of parameters falling above the straight lines on the six half-normal plots. For I\textsubscript{Kto} $\{p_{1},\ p_{2},\ p_{3},\ p_{4},\ p_{5},\ p_{7},\ p_{11}\}$ are selected, for I\textsubscript{Kslow1} $\{p_{1},\ p_{2},\ p_{4},\ p_{5},\ p_{9},\ p_{10}\}$, for I\textsubscript{Kslow2} $\{p_{1},\ p_{2}\}$, and for I\textsubscript{Kss} $\{p_{1},\ p_{2},\ p_{3}\}$. 
\begin{figure}
    \centering
    \includegraphics{figs/half-norm.pdf}
    \caption{Half-normal plots of factorial effects of parameters on the six maker points.}
    \label{fig:half-norm}
\end{figure}
\textbf{Figure~\ref{fig:calib_param}} shows the selected parameters, highlighted in different colors according to their functional roles in channel kinetics. We categorized the calibration parameters into four classes: The red represents the voltage-threshold parameters and the green voltage slopes, controlling the voltage dependence, the blue scale factors of kinetic functions, and purple time-constant shifters. Note that the voltage-dependence parameters in red and green appear multiple times across different equations. This parsimonious model design is intended to maximize the structural regularization to minimize overfitting. In general, voltage-threshold and time-constant-shifting parameters impact the current traces more than others. 
\begin{figure}[!ht]
    \centering
    \includegraphics{figs/calib_param.pdf}
    \caption{Selected parameters from the sensitivity analysis, highlighted in different colors according to their functional roles in channel kinetics.}
    \label{fig:calib_param}
\end{figure}

\subsection{Model Fitness}
We used the BFGS algorithm, one of the most popular nonlinear optimization methods, in the proposed model calibration routine because it supports box constraints.

box-constrained BFGS / goodness-of-fit measure / problem of rmse (not normalized) / problem of classical r2 / so, we chose a new r2 measure developed by UCLA guys.

We compared three nonlinear optimization algorithms satisfying box constraints: BFGS, sequential quadratic programming (SQP), and active set methods. \textbf{Fig~\ref{fig:model_fitness}A} and \textbf{C} shows values of the objective function (sum of RMSEs) of the three optimization algorithms for WT and MGAT1KO, respectively. The three algorithms barely showed differences in the objective values. We compared the actual model predictions and experimental potassium I\textsubscript{K} traces in \textbf{Figure~\ref{fig:model_fitness}B} and \textbf{D} to validate the model calibration results further. Representative data were selected randomly for each group and plotted at -30 mV and 50 mV voltage steps to illustrate discrepancies at low and high voltages. Sum of RMSEs values are lower in MGAT1KO than WT in general, because magnitudes are smaller for the disease group. Calibration results of the BFGS algorithm will be used in analysis of kinetics modeling in the rest the sections. 
\begin{figure}[!ht]
    \centering
    \includegraphics{figs/model_fitness_4half.pdf}
    \caption{Bar graphs of the sum of RMSEs of (A) WT and (C) MGAT1KO, and representative actual fitness between model outputs and experimental I\textsubscript{K} recordings (in red) at -30 mV and 50 mV voltage steps of (B) observation 26 in WT and (D) 4 in MGAT1KO.}
    \label{fig:model_fitness}
\end{figure}
